---
title: "Portfolio 1: Part 1.4 LENA KOGAN"
output: html_notebook
---
Part 1: Maps
```{r}
library("BIEN")
library(ape) #Package for working with phylogenies in R
library(maps) #Useful for making quick maps of occurrences
library(sp) # A package for spatial data
library(dplyr)
library(lattice)
```

Now, I am making the map.This species does not have any range data in BIEN, so I have made a map with only the occurence data.
```{r}
#a_arabicum_range <- BIEN_ranges_load_species("Aethionema arabicum")
a_arabicum_occurence <- BIEN_occurrence_species(species = "Aethionema arabicum")
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
#plot(a_arabicum_range, col="green", add=T)
points(cbind(a_arabicum_occurence$longitude, a_arabicum_occurence$latitude), col="blue", pch=20, cex=1)
```
Now I am making a map for the avocado plant.
```{r}
avocado_range <- BIEN_ranges_load_species("Persea americana")
avocado_occurence <- BIEN_occurrence_species(species = "Persea americana")
map('world', fill=T, col="grey", bg="light blue", xlim=c(-180,-20), ylim=c(-60, 80))
plot(avocado_range, col="green", add=T)
points(cbind(avocado_occurence$longitude, avocado_occurence$latitude), col="blue", pch=20, cex=1)
```
PART 2: TRAIT DATA

Species 1: Avocados
I am comparing two species, Persea americana and Persea borbonia. First, I will do Persea americana,
```{r}
Persea_americana <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Persea americana")
str(Persea_americana)
Persea_americana
```

Now, I am going to do Persea borbonia.
```{r}
Persea_borbonia <- BIEN_trait_traitbyspecies(trait = "whole plant height", species = "Persea borbonia")
str(Persea_borbonia)
Persea_borbonia
```
Now, I will commute the mean of each of these avocado species.
```{r}
americana = Persea_americana[c(3)]
americana = as.data.frame(americana)
americana_height = as.numeric(americana$trait_value)
mean(americana_height, data=americana)
borbonia = Persea_borbonia[c(3)]
borbonia = as.data.frame(borbonia)
borbonia_height = as.numeric(borbonia$trait_value)
mean(borbonia_height, data=borbonia)
```
```{r}
boxplot(americana_height)
```

```{r}
boxplot(borbonia_height)
```
```{r}
t.test(americana_height, borbonia_height)
```

Because the p-value is less than 0.05, (in fact, it's an extremely small value), the height difference between the two species of avocados is significant.

Species 2: Cashew
I am going to compare two species of cashew I am comparing Anacardium occidentale and Anacardium excelsum.
```{r}
Anacardium_occidentale <- BIEN_trait_traitbyspecies(trait = "seed mass", species = "Anacardium occidentale")
str(Anacardium_occidentale)
Anacardium_occidentale
```
```{r}
Anacardium_excelsum <- BIEN_trait_traitbyspecies(trait = "seed mass", species = "Anacardium excelsum")
str(Anacardium_excelsum)
Anacardium_excelsum
```
```{r}
excelsum = Anacardium_excelsum[c(3)]
excelsum = as.data.frame(excelsum)
excelsum_mass = as.numeric(excelsum$trait_value)
mean(excelsum_mass, data=excelsum)
occidentale = Anacardium_occidentale[c(3)]
occidentale = as.data.frame(occidentale)
occidentale_mass = as.numeric(occidentale$trait_value)
mean(occidentale_mass, data=borbonia)
```
```{r}
boxplot(occidentale_mass)
```
```{r}
boxplot(excelsum_mass)
```

```{r}
t.test(excelsum_mass, occidentale_mass)

```




